---
title: "Submit2"
author: "Runan Yao"
date: "October 2, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
```

### Read data
```{r}
allDegResult <- read.csv("Phase_1_Result/FC FDR of each cell line/AllDEGResult.csv", stringsAsFactors = FALSE)
colnames(allDegResult)
```



### Preprocess
1. Regroup
```{r}
allNames <- colnames(allDegResult)
expGroupNames <- c( 'XPC3TR.XPC3NT___log2FoldChange', 
                    'XPC3TR.XPC3NT___padj',
                    'XPC3MTR.XPC3MNT___log2FoldChange',
                    'XPC3MTR.XPC3MNT___padj',
                    'XMCF7TR.XMCF7NT___log2FoldChange',
                    'XMCF7TR.XMCF7NT___padj')
ExpGroup <- allDegResult %>% 
  select(one_of( c( 'Gene',expGroupNames)))

colnames(ExpGroup) <- c('Gene', 
                        'Exp_LFC_1', 'Exp_Alpha_1', 
                        'Exp_LFC_2', 'Exp_Alpha_2', 
                        'Exp_LFC_3', 'Exp_Alpha_3')
sampleNames.Exp <- c('PC3','PC3M','MCF7')

CtrlGroup <- allDegResult %>%
  select(one_of(setdiff( allNames, c(expGroupNames, c('X', 'Symbol_184A1', 'XTRAMPTR.XTRAMPNT___log2FoldChange', 'XTRAMPTR.XTRAMPNT___padj')))))

colnames(CtrlGroup) <- c('Gene', 
                         'CTRL_LFC_1', 'CTRL_Alpha_1', 
                         'CTRL_LFC_2', 'CTRL_Alpha_2', 
                         'CTRL_LFC_3', 'CTRL_Alpha_3',
                         'CTRL_LFC_4', 'CTRL_Alpha_4',
                         'CTRL_LFC_5', 'CTRL_Alpha_5',
                         'CTRL_LFC_6', 'CTRL_Alpha_6',
                         'CTRL_LFC_7', 'CTRL_Alpha_7',
                         'CTRL_LFC_8', 'CTRL_Alpha_8')
sampleNames.Ctrl <- c('184A1', 'C32', 'CCD18', 'HEM', 'HT29', 'hTERT', 'PANC', 'RW') 

```

2. In each cell-line, for each gene, if adj-alpha value is < threshold, then keep LFC otherwise set LFC = 0
```{r}
ourThreshold <- 0.5
keepOrNot <- function(alpha, threshold){
  return(
    ifelse(is.na(alpha), 
           0, 
           ifelse(alpha < threshold, 
                  1, 
                  0)
           )
    )
}

Exp1 <- ExpGroup$Exp_LFC_1 * keepOrNot(ExpGroup$Exp_Alpha_1, threshold = ourThreshold)
Exp2 <- ExpGroup$Exp_LFC_2 * keepOrNot(ExpGroup$Exp_Alpha_2, threshold = ourThreshold)
Exp3 <- ExpGroup$Exp_LFC_3 * keepOrNot(ExpGroup$Exp_Alpha_3, threshold = ourThreshold)

Ctrl1 <- CtrlGroup$CTRL_LFC_1 * keepOrNot(CtrlGroup$CTRL_Alpha_1, threshold = ourThreshold)
Ctrl2 <- CtrlGroup$CTRL_LFC_2 * keepOrNot(CtrlGroup$CTRL_Alpha_2, threshold = ourThreshold)
Ctrl3 <- CtrlGroup$CTRL_LFC_3 * keepOrNot(CtrlGroup$CTRL_Alpha_3, threshold = ourThreshold)
Ctrl4 <- CtrlGroup$CTRL_LFC_4 * keepOrNot(CtrlGroup$CTRL_Alpha_4, threshold = ourThreshold)
Ctrl5 <- CtrlGroup$CTRL_LFC_5 * keepOrNot(CtrlGroup$CTRL_Alpha_5, threshold = ourThreshold)
Ctrl6 <- CtrlGroup$CTRL_LFC_6 * keepOrNot(CtrlGroup$CTRL_Alpha_6, threshold = ourThreshold)
Ctrl7 <- CtrlGroup$CTRL_LFC_7 * keepOrNot(CtrlGroup$CTRL_Alpha_7, threshold = ourThreshold)
Ctrl8 <- CtrlGroup$CTRL_LFC_8 * keepOrNot(CtrlGroup$CTRL_Alpha_8, threshold = ourThreshold)

length(Exp1)
LFCTable <- cbind(Exp1,Exp2,Exp3,Ctrl1,Ctrl2,Ctrl3,Ctrl4,Ctrl5,Ctrl6,Ctrl7,Ctrl8) 

dim(LFCTable)
colnames(LFCTable) <- c('Exp_1','Exp_2','Exp_3','Ctrl_1','Ctrl_2','Ctrl_3','Ctrl_4','Ctrl_5','Ctrl_6','Ctrl_7','Ctrl_8')
rownames(LFCTable) <- ExpGroup$Gene

LFCTable[is.na(LFCTable)] <- 0

# 
# upRegLFCTable <- ifelse(LFCTable > 0, LFCTable, 0 )
# ConvertedUpRegLFCTable <- floor(2^upRegLFCTable * 1000)
# 

NormalizeLFCNumber <- function(x){
  return(x-min(x))
}

normalizedLFCTable <- apply(LFCTable,2, NormalizeLFCNumber)
summary(normalizedLFCTable)

convertedLFCTable <- normalizedLFCTable
summary(convertedLFCTable)

write.csv(LFCTable, file='Phase_2_Result/Filtered_LFC_Result.csv', row.names = TRUE)
write.csv(convertedLFCTable, file='Phase_2_Result/convertedLFC_Result.csv', row.names = TRUE)
LFCTable <- as.data.frame(LFCTable, stringsAsFactors = FALSE)

```


### all zero LFC
```{r}
dim(LFCTable)
result <- apply(LFCTable,1, function(x){all(x==0)})

sum(result)
```

## 

For each gene, assume the gene LFC can only be picked in one of two ditributions. Then the chance of having Exp_1~Exp_3 from one group and Ctrl_1~Ctrl_8 from the other group is $\frac{2}{2^{11}}$. 

### Different direction of express in between groups
The ideal case is:

1.  Exp group all up, Ctrl group all down; or 

2.  Exp group all down, Ctrl group all up

```{r}
sameInExpUpCtrlDownOrKeep <- apply(LFCTable,1, function(x){ all(x[1:3] > 0) && all(x[4:11] <= 0) } )
sameInExpUpOrKeepCtrlDown <- apply(LFCTable,1, function(x){ all(x[1:3] >= 0) && all(x[4:11] < 0) } )
sameInExpDownCtrlUpOrKeep <- apply(LFCTable,1, function(x){ all(x[1:3] < 0) && all(x[4:11] >= 0) } )
sameInExpDownOrKeepCtrlUp <- apply(LFCTable,1, function(x){ all(x[1:3] <= 0) && all(x[4:11] > 0) } )

names(sameInExpUpOrKeepCtrlDown) = NULL
names(sameInExpUpCtrlDownOrKeep) = NULL
names(sameInExpDownCtrlUpOrKeep) = NULL
names(sameInExpDownOrKeepCtrlUp) = NULL

which(sameInExpUpOrKeepCtrlDown == TRUE) 
which(sameInExpUpCtrlDownOrKeep == TRUE)
which(sameInExpDownCtrlUpOrKeep == TRUE) 
which(sameInExpDownOrKeepCtrlUp == TRUE) 
```


Here we can see, none of those genes follow the ideal case.

### T.test

```{r}
matrixLFC <- as.matrix(LFCTable)
colnames(matrixLFC) <- NULL
rownames(matrixLFC) <- NULL
runTTest <- function(x){
  return(t.test(x[1:3], x[4:11] )$p.value)
}
tTestResult <- apply(matrixLFC, 1, runTTest)
sum(tTestResult<0.05, na.rm = TRUE)

length(tTestResult)
ExpGroup$Gene[which(tTestResult<0.05)]


```



```{r}





```